A multi-dimensional extension of the Lightweight Temporal Compression method

### 2018-10-17 paper.tex
Some corrections

\author{Bo Li$^1$, Omid Sarbishei$^2$, Tristan Glatard$^1$\\
  $^1$ Department of Computer Science and Software Engineering, Concordia University, Montreal, QC, Canada \\
  $^2$ Motsai, Saint Bruno, QC, Canada.}
\maketitle

@@

\todo{In all of our experiments, the board was attached to the lower 
right arm as shown in Fig. X. (Put a picture of Neblina in a box worn 
on the wrist)}

In this experiment, we validated the behavior of our LTC extension on a 
PC using data acquired with Neblina. We collected two 3D 
accelerometer time series, a short one and a longer one, acquired 
on two different subjects performing biceps curl, with a 50Hz sampling rate (see 
Figure~\ref{fig:datasets-1}). It should be noted that the longer dataset also has 
a higher amplitude, most likely due to differences between subjects.

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-x.pdf}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-y.pdf}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-z.pdf}
\caption{Short biceps curl}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-x.pdf}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-y.pdf}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-z.pdf}
\caption{Long biceps curl} 
\end{subfigure}
\caption{Time series used in Experiment 1}
\label{fig:datasets-1}
\end{figure*}

@@

\paragraph{Inifinity vs Euclidean norms}
Compared to the infinity norm, the average ratio between the compression ratios obtained
with the infinity and Euclidean norm is 1.08 for 2D data, and 1.20 
for 3D data. These ratios are lower than the theoretical values of 
$\frac{4}{\pi}$ in 2D and $\frac{6}{\pi}$ in 3D, which are obtained for 
random-uniform signals. Unsurprisingly, the infinity norm surpasses the 
Euclidean norm in terms of resource consumption. Memory-wise, the 
infinity norm requires a constant amount of 80~B, used to store the 
intersection of n-balls. The Euclidean norm, however, uses up to 4.9~KB of memory 
for the Long time series in 3D with $\epsilon=100$. More importantly, 
the amount of required memory increases for more complex time series 
(Long vs Short), and it also increases with larger 
values of $\epsilon$. Similar observations are made for the processing 
time, with values ranging from 0.2~ms for the simplest time series and 
smallest $\epsilon$, to 26.4~ms for the most complex time series and 
largest $\epsilon$.

@@

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-x.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-y.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-z.pdf}
\caption{Walking}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-x.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-y.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-z.pdf}
\caption{Running}
\end{subfigure}
\caption{Time series used in Experiment 2}
\label{fig:datasets-2}
\end{figure*}
\subsection{Experiment 2: impact on energy consumption}

@@

The choice of an appropriate norm should not be underestimated. This problem 
is specific to multi-dimensional streams, since all common norms are 
identical in 1D. Some situations might be better described with the 
Euclidean norm than with the infinity norm, such as the ones involving 
position or movement measures. Using the infinity norm instead of the 
Euclidean would lead to important error differences, proportional to 
$\sqrt{n}$ in dimension $n$. A better algorithm to detect the 
intersection of n-balls in our context would improve the efficiency of LTC
in dimension $n$ for arbitrary norms.

>>>>

\author{Bo Li$^1$, Omid Sarbishei$^2$, Tristan Glatard$^1$\\
  $^1$ Department of Computer Science and Software Engineering, Concordia University, Montreal, QC, Canada \\
  $^2$  Research and Development Department, Motsai Research, Saint Bruno, QC, Canada}
\maketitle

@@

\todo{(Put a picture of Neblina in a box worn 
on the wrist)}

In this experiment, we validated the behavior of our LTC extension on a 
PC using data acquired with Neblina. We collected two 3D 
accelerometer time series, a short one and a longer one, acquired 
on two different subjects performing biceps curl, with a 50Hz sampling rate (see 
Figure~\ref{fig:datasets-1}). In both cases, the 
subject was wearing Neblina on their right wrist. It should be noted that the longer dataset also has 
a higher amplitude, most likely due to differences between subjects.

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-x.pdf}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-y.pdf}
\includegraphics[width=0.33\columnwidth]{figures/5-time-bicep-curl-plot-z.pdf}
\caption{Short biceps curl$^a$}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-x.pdf}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-y.pdf}
\includegraphics[width=0.33\columnwidth]{figures/Mohammad-bicep-curl-plot-z.pdf}
\caption{Long biceps curl} 
\end{subfigure}
\caption{Time series used in Experiment 1\\
{\footnotesize $^a$ Average of az data is -7.81g. It was shifted to 0 so that the graphs can all use the same y scale.}}
\label{fig:datasets-1}
\end{figure*}

@@

\paragraph{Inifinity vs Euclidean norms}
Compared to the infinity norm, the average ratio between the compression ratios obtained
with the infinity and Euclidean norm is 1.08 for 2D data, and 1.20 
for 3D data. These ratios are lower than the theoretical values of 
$\frac{4}{\pi}$ in 2D and $\frac{6}{\pi}$ in 3D, which are obtained for 
random-uniform signals. Unsurprisingly, the infinity norm surpasses the 
Euclidean norm in terms of resource consumption. Memory-wise, the 
infinity norm requires a constant amount of 80~B, used to store the 
intersection of n-balls. The Euclidean norm, however, uses up to 4.9~KB of memory 
for the Long time series in 3D with $\epsilon=100$. More importantly, 
the amount of required memory increases for more complex time series 
(Long vs Short), and it also increases with larger 
values of $\epsilon$. Similar observations are made for the processing 
time, with values ranging from 0.2~ms for the simplest time series and 
smallest $\epsilon$, to 26.4~ms for the most complex time series and 
largest $\epsilon$. Figure~\ref{fig:memory} shows the memory consumption 
of the 3D Euclidean implementation for $\epsilon=48.8~mg$: peaks appear
at the end of long compressed ranges where the signal was $\epsilon$-closed
to the linear approximation.

@@

\begin{figure*}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-x.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-y.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/walking-1000-z.pdf}
\caption{Walking}
\end{subfigure}
\begin{subfigure}{2\columnwidth}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-x.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-y.pdf}
\includegraphics[width=0.33\columnwidth]{./figures/running-1000-z.pdf}
\caption{Running}
\end{subfigure}
\caption{Time series used in Experiment 2}
\label{fig:datasets-2}
\end{figure*}

\begin{figure}
\includegraphics[width=\columnwidth]{./figures/memory-mohammad.pdf}
\caption{Memory usage of 3D Euclidean implementation, long biceps curl time series, $\epsilon$=48.8~mg.}
\label{fig:memory}
\end{figure}

\subsection{Experiment 2: impact on energy consumption}

@@

The choice of an appropriate norm should not be underestimated. This problem 
is specific to multi-dimensional streams, since all common norms are 
identical in 1D. Some situations might be better described with the 
Euclidean norm than with the infinity norm, such as the ones involving 
position or movement measures. Using the infinity norm instead of the 
Euclidean would lead to important error differences, proportional to 
$\sqrt{n}$ in dimension $n$. A better algorithm to detect the 
intersection of n-balls in our context would improve the efficiency of 
LTC in dimension $n$ for arbitrary norms. Investigating other norms, in 
particular the 1-norm, would be relevant too.
